---
title: "LAGOS Spatial Analysis"
author: "Laney Gordon"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: html_document
editor_options: 
  chunk_output_type: console
---



```{r setup, include=FALSE}
library(tidyverse) # Tidy packages
library(sf) #Spatial package that can read and create shapefiles 
library(mapview) #Interactive maps
  mapviewOptions(fgb= FALSE) 
library(LAGOSNE) #Lots and lots of clean lake data
library(USAboundaries) #USA states and counties
```
```{r}
#install.packages(c("RApiSerialize", "LAGOSNE", 'USAboundaries'))

LAGOSNE::lagosne_get(dest_folder = LAGOSNE:::lagos_path()) 
```
# LAGOS Analysis


## Loading in data


### First download and then specifically grab the locus (or site lat longs)

```{r data-read}
# #Lagos download script
LAGOSNE::lagosne_get(dest_folder = LAGOSNE:::lagos_path())

#Load in lagos
lagos <- lagosne_load()

#Grab the lake centroid info
lake_centers <- lagos$locus



```



### Convert to spatial data
```{r}

spatial_lakes <- st_as_sf(lake_centers,coords=c('nhd_long','nhd_lat'),
                          crs=4326) %>%
  st_transform(2163)

subset_spatial <- spatial_lakes %>%
  slice(1:100) 

subset_baser <- spatial_lakes[1:100,]

#Dynamic mapviewer
mapview(subset_spatial)

```


### Subset to only Minnesota

```{r}
states <- us_states()

#Plot all the states to check if they loaded
#mapview(states)
minnesota <- states %>%
  filter(name == 'Minnesota') %>%
  st_transform(2163)

#Subset lakes based on spatial position
minnesota_lakes <- spatial_lakes[minnesota,]

#Plotting the first 1000 lakes
minnesota_lakes %>%
  arrange(-lake_area_ha) %>%
    slice(1:1000) %>%
  mapview(.,zcol = 'lake_area_ha')
```


# In-Class work


## 1) Show a map outline of Iowa and Illinois (similar to Minnesota map upstream)


```{r}
#Iowa and Illinois Lakes
spatial_lakes <- st_as_sf(lake_centers,coords=c('nhd_long','nhd_lat'),
                          crs=4326) %>%
  st_transform(2163)

subset_spatial <- spatial_lakes %>%
  slice(1:100) 

subset_baser <- spatial_lakes[1:100,]


states <- us_states()

iowa_il <- states %>%
  filter(name %in% c('Iowa', 'Illinois')) %>%
  st_transform(2163)

mapview(iowa_il) 


```



## 2) Subset LAGOS data to these sites, how many sites are in Illinois and Iowa
combined? How does this compare to Minnesota?

IA + IL sites = 16,466 obs of 17 variables
MN sites = 29,038 obs of 17 variables
If we are comparing data sites, Minnesota has more than Illinois and Iowa combined.

```{r}
#subset using LAGOS data

iowa_il_lakes <- spatial_lakes[iowa_il,] 


```


## 3) What is the distribution of lake size in Iowa vs. Minnesota?

- Here I want to see a histogram plot with lake size on x-axis and frequency on 
y axis (check out geom_histogram)

```{r} 
#Iowa lakes
spatial_lakes <- st_as_sf(lake_centers,coords=c('nhd_long','nhd_lat'),
                          crs=4326) %>%
  st_transform(2163) 

subset_spatial <- spatial_lakes %>%
  slice(1:100) 

subset_baser <- spatial_lakes[1:100,]

states <- us_states()
iowa<- states %>%
  filter(name == "Iowa") %>%
  st_transform(2163) 

iowa_lakes <- spatial_lakes[iowa,]



#Minnesota Lakes
minnesota <- states %>%
  filter(name == 'Minnesota') %>%
  st_transform(2163) 
 

minnesota_lakes <- spatial_lakes[minnesota,] 



iowa_MN <- states %>%
  filter(name %in% c('Iowa', 'Minnesota')) %>% 
   st_transform(2163)  
  
IA_MN_lakes <- spatial_lakes %>% 
  .[iowa_MN,] %>%
  st_join(iowa_MN) 


```
```{r}

#IA vs MN Lake Area Histogram 

ggplot() +
  geom_histogram(filter(IA_MN_lakes, name == "Minnesota"), mapping = aes(lake_area_ha), bins= 15, color="darkblue", fill="lightblue")+
  scale_x_log10() +
  labs(title= "Iowa vs Minnesota Lake Size", x= "Lake Area (ha)", y= 'Frequency') + 

geom_histogram(filter(IA_MN_lakes, name == "Iowa"), mapping = aes(lake_area_ha), bins = 15, color="red", fill="pink")+
  scale_x_log10() +
  labs(title="Iowa vs Minnesota Lake Size", x="Lake Area(ha)", y='Frequency')+
  
 facet_wrap(~name) 

```


## 4) Make an interactive plot of lakes in Iowa and Illinois and color them
by lake area in hectares

```{r}
iowa_il_lakes %>%
  arrange(-lake_area_ha) %>%
    slice(1:1000) %>%
  mapview(.,zcol = 'lake_area_ha') 

```


## 5) What other data sources might we use to understand how reservoirs and 
natural lakes vary in size in these three states? 

There could be copious accounts of scientific research that would help us understand how lakes vary in size throughout Minnesota, Iowa, and Illinois. Climate and ecological data surrounding these lakes would allow for further analysis of possible patterns and differences from state to state. The following research was conducted by the Minnesota Pollution Control Agency as a report for the EPA's 2007 National Lake Assessment [National Lake Assessment](https://www.pca.state.mn.us/sites/default/files/wq-nlap1-05.pdf)       
