---
title: "Snow Data Assignment: Web Scraping, Functions, and Iteration"
author: "Laney Gordon"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
library(rvest)
library(tidyverse)
library(lubridate)
library(readxl)

```


# Assignment:

1. Extract the meteorological data URLs. Here we want you to use the `rvest` package to get the URLs for the `SASP forcing` and `SBSP_forcing` meteorological datasets.

```{r, message=F}
site_url <- 'https://snowstudies.org/archived-data/'

webpage <- read_html(site_url)

links <- webpage %>%
  html_nodes('a') %>%
  .[grepl('forcing',.)] %>%
  html_attr('href')


```


2. Download the meteorological data. Use the `download_file` and `str_split_fixed` commands to download the data and save it in your data folder. You can use a for loop or a map function. 

```{r, message= F}
splits <- str_split_fixed(links,'/',8)

dataset<- splits[,8]

Q2_file_names <- paste0('data/',dataset)

for(i in 1:2){
  download.file(links[i],destfile=Q2_file_names[i])
}

downloaded <- file.exists(Q2_file_names)

evaluate <- !all(downloaded)


```

3. Write a custom function to read in the data and append a site column to the data. 

```{r, message=F}

library(pdftools)
headers <- pdf_text('https://snowstudies.org/wp-content/uploads/2022/02/Serially-Complete-Metadata-text08.pdf') %>%
  readr::read_lines(.) %>%
  trimws(.) %>%
  str_split_fixed(.,'\\.',2) %>%
  .[,2] %>%
  .[1:26] %>%
  str_trim(side = "left")

```



```{r, message=F}
our_meteor_reader <- function(Q3_file){
  name = str_split_fixed(Q3_file,'/',2)[,2] 
  name2 = str_split_fixed(Q3_file,'/',4)[,2] 
  df <- read.delim(Q3_file,header = FALSE, sep = "", col.names = headers, skip = 4) %>% 
    select(1:14) %>%
    mutate(site = name2) 

} 
  
```

4. Use the `map` function to read in both meteorological files. Display a summary of your tibble.

```{r}
meteor_data_full <- map_dfr(Q2_file_names,our_meteor_reader)

summary(meteor_data_full)

```


5. Make a line plot of mean temp by year by site (using the `air temp [K]` variable). Is there anything suspicious in the plot? Adjust your filtering if needed.

```{r,message=F, fig.cap="Figure 1: Lines of the graph represent a time series of average monthly temperature in Kelvin at the Swamp Angel Study Plot and the Senator Beck Study Plot. The year 2003 was not recorded due to incomplete datasets."}
temp_yearly <- meteor_data_full %>%
  filter(year > 2003) %>%
  group_by(year, site) %>%
  summarize(mean_temp = mean(air.temp..K., na.rm=T))

ggplot(temp_yearly, aes(x = year, y = mean_temp, color=site)) +
  geom_line() +
  ggthemes::theme_few() +
  ggthemes::scale_color_few() +
  labs(x = "Year",
       y = "Average Temp (K)", 
     color="Site") +
  scale_color_manual(labels = c("Swamp Angel", "Senator Beck"), values=c("green", "pink")) +
     theme(legend.position = "bottom")

```

6. Write a function that makes line plots of monthly average temperature at each site for a given year. Use a for loop to make these plots for 2005 to 2010. Are monthly average temperatures at the Senator Beck Study Plot ever warmer than the Snow Angel Study Plot?
Hint: https://ggplot2.tidyverse.org/reference/print.ggplot.html

Based on the plots, it appears that the largest differences in temperature occurs during the summer months and the smallest change occurs in fall months. Although, the Senator Beck Study plot is never warmer than the Swamp Angel Study Plot. 
```{r, message=F}
lineplotter <- function(meteor_data_full, year) {
  temp_monthly <- meteor_data_full %>% 
    group_by(month, year, site) %>%
  summarize(mean_temp = mean(air.temp..K.)) %>%
    filter (yr == year)
  

  plots2 <- ggplot(temp_monthly, aes(x = month, y= mean_temp, color= site)) +
    geom_line() +
    labs(x = "Month", 
         y= "Average Temp (K)",
         title= temp_monthly$year,
         color= "Site") +
    scale_color_manual(labels = c("Swamp Angel", "Senator Beck"), values= c("green", "pink"))
  print(plots2)
}

years <- c(2005, 2006, 2007, 2008, 2009, 2010)
for (yr in years) {
  lineplotter(meteor_data_full, year)
}
```


Bonus: Make a plot of average daily precipitation by day of year (averaged across all available years). Color each site. 

Bonus #2: Use a function and for loop to create yearly plots of precipitation by day of year. Color each site. 
