---
title: "Lake Water Quality Analysis"
author: "Laney Gordon"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: html_document
---



```{r setup, include=FALSE}
library(tidyverse) # Tidy packages
library(sf) #Spatial package that can read and create shapefiles 
library(mapview) #Interactive maps
library(LAGOSNE) #Lots and lots of clean lake data
library(USAboundaries) #USA states and counties
library(lubridate) #For dealing with date and time
```


# LAGOS Analysis


## Loading in data


### First download and then specifically grab the locus (or site lat longs)
```{r data-read}
#Lagos download script
#lagosne_get(dest_folder = LAGOSNE:::lagos_path(),overwrite=T)

#Load in lagos
lagos <- lagosne_load()


#Grab the lake centroid info
lake_centers <- lagos$locus

# Make an sf object 
spatial_lakes <- st_as_sf(lake_centers,coords=c('nhd_long','nhd_lat'),
                          crs=4326)

#Grab the water quality data
nutr <- lagos$epi_nutr

#Look at column names
#names(nutr)
```

### Subset columns nutr to only keep key info that we want


```{r}
clarity_only <- nutr %>%
  select(lagoslakeid,sampledate,chla,doc,secchi) %>%
  mutate(sampledate = as.character(sampledate) %>% ymd(.))

```


### Keep sites with at least 200 observations 

```{r}

#Look at the number of rows of dataset
#nrow(clarity_only)

chla_secchi <- clarity_only %>%
  filter(!is.na(chla),
         !is.na(secchi))

# How many observatiosn did we lose?
# nrow(clarity_only) - nrow(chla_secchi)


# Keep only the lakes with at least 200 observations of secchi and chla
chla_secchi_200 <- chla_secchi %>%
  group_by(lagoslakeid) %>%
  mutate(count = n()) %>%
  filter(count > 200)


```


### Join water quality data to spatial data

```{r}
spatial_200 <- inner_join(spatial_lakes,chla_secchi_200 %>%
                            distinct(lagoslakeid,.keep_all=T),
                          by='lagoslakeid')


```

### Mean Chl_a map

```{r}
### Take the mean chl_a and secchi by lake

mean_values_200 <- chla_secchi_200 %>%
  # Take summary by lake id
  group_by(lagoslakeid) %>%
  # take mean chl_a per lake id
  summarize(mean_chl = mean(chla,na.rm=T),
            mean_secchi=mean(secchi,na.rm=T)) %>%
  #Get rid of NAs
  filter(!is.na(mean_chl),
         !is.na(mean_secchi)) %>%
  # Take the log base 10 of the mean_chl
  mutate(log10_mean_chl = log10(mean_chl)) %>%
  # Take the log base 10 of the mean_secchi
  mutate(log10_mean_secchi = log10(mean_secchi))

#Join datasets
mean_spatial <- inner_join(spatial_lakes,mean_values_200,
                          by='lagoslakeid') 

#Make a map
mapview(mean_spatial,zcol='log10_mean_chl')

```


# Class work

## 1) What is the correlation between Secchi Disk Depth and Chlorophyll a for
sites with at least 200 observations?

- Here, I just want a plot of chla vs secchi for all sites 

```{r, warning=FALSE}
ggplot(chla_secchi_200, aes(y= chla, x= secchi))+
  geom_point(size=1, shape = 1) +
  scale_x_log10() +
  scale_y_log10() 


```

## Why might this be the case? 
Based on the map, it appears that if there is more chlorophyll or other nutrients in the water, it makes it harder to read the Secchi Disk. 


## 2) What states have the most data? 
Minnesota

### 2a) First you will need to make a lagos spatial dataset that has the total 
number of counts per site.

```{r}

site_counts <- nutr %>%
  group_by(lagoslakeid) %>%
  summarize(count = n()) 

spacial_counts <- inner_join(spatial_lakes, site_counts %>%
                               distinct(lagoslakeid,.keep_all=TRUE),
                             by= "lagoslakeid") 

``` 


### 2b) Second, you will need to join this point dataset to the us_boundaries 
data. 

```{r}

states_lagos <- lagos$state %>%
  select(-state_pct_in_nwi, -state_ha_in_nwi,-state_ha) 

states_lakes <- inner_join(spacial_counts, states_lagos, by = "state_zoneid")

```

 


### 2c) Then you will want to group by state and sum all the observations in that
state and arrange that data from most to least total observations per state. 

```{r}
sum_stats <- states_lakes %>%
  group_by(state_name) %>%
  summarize(sum_counts = sum(count)) %>%
  as.data.frame() %>% 
  select(-geometry) %>% 
  arrange(desc(sum_counts))

head(sum_stats) 
```

Minnesota has the most data. This phenomenon most likely occurs because this state has more lakes than any other state, and reporting data may be beneficial for agricultural purposes, such as irrigation.

##3 Is there a spatial pattern in Secchi disk depth for lakes with at least 200 
observations?

```{r}

mapview(mean_spatial,zcol='log10_mean_secchi') 
```

After creating a color coordinated map view of the Secchi disk depth of different lakes across the north east, we conclude that there is not a clear spatial pattern. Lakes are clustered near Minneapolis and the majority of Rhode Island, although these two locations do not appear to have any other significant factors in common.      



